---
title: "BEC_Database"
author: "Desira Miller"
date: '2022-06-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r load-packages, include = FALSE}
# install.packages("tindyverse")
# install.packages("readxl")
library("readxl")
library("tidyverse")
library("dplyr")
library("openintro")
library("skimr")
library("janitor")
library("psych")
library("stringr")
```


## R Markdown
## Explore Data, clean col names, convert to lowercase

### Add another line in your mutate function to calculate horizon depth >- 100 cm

```{r read-in-data, warning = F}
data_bec_database<- read_excel("8 BEC Database.xlsx")
glimpse(data_bec_database)
```

## Retain Tidy Excel Columns relevant col; create new dataframe

```{r new-dataframe-with-select-variables}
``` 

## Create an index column specifiying if a bulk density value is gap-filled or not

Create a vector of gap-filled values

```{r bulk-density-gap-filled-values}
```

Create the `bd_fill` index column

```{r create-gap-filled-index}

```

## Extract soil horizon from `horizon` and call it `horizon_simple`
```{r create-horizon-simple}

```
## Calculate carbon content for each horizon

```{r mutate-cf-na-to-0}

```

```{r calculate-ccontent-from-function}

```
## Create carbon_stock and carbon_stock_1M variables
```{r create-carbon_stock-variable}

```
```{r create-carbon_stock_1M-variable}

```




## Save the tidy dataset as a .csv

```{r write-csv}

```

## Create summary tables for `carbon_stock`, `carbon_stock_1m`

```{r carbon_stock-summary-table}

```

```{r carbon_stock_1m-summary-table}

```


