---
title: "Seak_NRCS_Final"
output: html_document
date: '2022-06-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r load-packages, include = FALSE}
# install.packages("tidyverse")
# install.packages("readxl")
library("readxl")
library("tidyverse")
library("dplyr")
library("openintro")
library("skimr")
library("janitor")
library("psych")
library("stringr")
```


## R Markdown
## Explore data , clean col names, convert to lowercase

### Add another line in your mutate function to calculate horizon depth >- 100 cm

```{r read-in-seak-nrcs-final-data, warning = F}
seak_nrcs_data <- read_excel("6 Seak NRCS FINAL.xlsx") %>% 
  janitor::clean_names() %>% 
  rename_with(tolower)
  
seak_nrcs_data
```

## Retain Tidy Excel Columns A-F,J-L, Q, S; create new dataframe

```{r new-dataframe-with-select-variables}

``` 

## Create an index column specifiying if a bulk density value is gap-filled or not

##Create a vector of gap-filled values

```{r bulk-density-gap-filled-values}

```

Create the `bd_fill` index column

```{r create-gap-filled-index}

```

## Extract soil horizon from `horizon` and call it `horizon_simple`
```{r create-horizon-simple}

```
## Calculate carbon content for each horizon

```{r mutate-cf-na-to-0}

```

```{r calculate-ccontent-from-function}

```


## Save the tidy dataset as a .csv

```{r}

```

## Create summary tables for `carbon_stock`, `carbon_stock_1m`

```{r carbon_stock-summary-table}

```

```{r carbon_stock_1m-summary-table}

```