---
title: "Shaw_Full_Profiles"
output: html_document
date: "2022-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
# install.packages("tindyverse")
# install.packages("readxl")
library("readxl")
library("tidyverse")
library("dplyr")
library("openintro")
library("skimr")
library("janitor")
library("psych")
library("stringr")
```


## R Markdown
## Explore Shaw Full Profiles, clean col names, convert to lowercase

### Add another line in your mutate function to calculate horizon depth >- 100 cm

```{r read-in-shaw-full-profiles-data, warning = F}
data <- read_excel("5 Shaw Full Profiles.xlsx") %>%  
  janitor::clean_names() %>% 
  rename_with(tolower)%>%
   mutate(bulk_density = as.numeric(bulk_density))
glimpse(data)
```

## Retain Tidy Excel Columns A-F,J-L, Q, S; create new dataframe
## Variables in this set are weird. Need to make a 'source' variable with every row being 'Shaw et. al 2018'.
## Need lat, lon, latlon_q, depth_1, depth_2, total_depth, depth_100, and cconc

```{r new-dataframe-with-select-variables, eval=T, echo=FALSE, message=FALSE, warning=FALSE}
data_subset <- data %>% 
  select(horizon_id,
         horizon, 
         upper_hzn_limit, 
         hzn_thickness,
         bulk_density,
         ccontent_1m)
glimpse(data_subset)
``` 

## Create an index column specifiying if a bulk density value is gap-filled or not

Create a vector of gap-filled values

```{r bulk-density-gap-filled-values, echo=FALSE}
bd_fill_values <- data_subset %>%
  select(bulk_density) %>%
  slice(1:3, 8:10, 16:28, 34:39, 46:146) %>%
  filter(!is.na(bulk_density)) %>% 
  distinct() %>% 
  pull() 
bd_fill_values
```

Create the `bd_fill` index column

```{r create-gap-filled-index, echo=FALSE}
data_subset <- data_subset %>% 
  mutate(bd_filled = ifelse(bulk_density %in% bd_fill_values, "yes", "no")) 
data_subset
```

## Extract soil horizon from `horizon` and call it `horizon_simple`
```{r create-horizon-simple, echo=FALSE}
data_subset$horizon_simple = substr(data_subset$horizon,1,1)
data_subset
```
## Calculate carbon content for each horizon

```{r mutate-cf-na-to-0, echo=FALSE, eval=FALSE}
data_subset <- data_subset %>%
 mutate(cf = ifelse(is.na(cf), 0, cf),
        cf_1_values = as.numeric(cf))
data_subset
```


## data_subset <- data_subset %>%
## mutate(ccontent = (cconc/100*bulk_density)*
##                  (100*100*total_depth)*
##                ((100 - cf)/100))
##data_subset


## Create carbon_stock and carbon_stock_1M variables
```{r create-shaw-full-profiles-carbon_stock-variable}
data_subset <- data_subset %>%
mutate(carbon_stock = (cconc*total_depth*bulk_density))
data_subset
```
```{r create-shaw-full-profiles-carbon_stock_1M-variable}
data_subset <- data_subset %>%
mutate(carbon_stock_1M = (ccontent_1m * depth_100 * bulk_density))
data_subset
```


## Save the tidy dataset as a .csv

```{r, echo=FALSE}
write.csv(data_subset, 'Shaw_Full_Profiles_Tidy.csv')
```

## Create summary tables for `carbon_stock`, `carbon_stock_1m`

```{r carbon_stock-summary-table, echo=FALSE}
describeBy(data_subset, group=data_subset$carbon_stock)
```

```{r carbon_stock_1m-summary-table, echo=FALSE}
describeBy(data_subset, group=data_subset$carbon_stock_1m)
```