---
title: "Shaw_Full_Profiles"
output: html_document
date: "2022-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
# install.packages("tidyverse")
# install.packages("readxl")
library("readxl")
library("tidyverse")
library("dplyr")
library("openintro")
library("skimr")
library("janitor")
library("psych")
library("stringr")
```


## R Markdown
## Explore Shaw Full Profiles, clean col names, convert to lowercase

### Add another line in your mutate function to calculate horizon depth >- 100 cm

```{r read-in-shaw-full-profiles-data, warning=F}
data <- read_excel("5 Shaw Full Profiles.xlsx") %>%  
  janitor::clean_names() %>% 
  rename_with(tolower)%>%
  mutate(bulk_density = as.numeric(bulk_density))
glimpse(data)
```

## Retain Tidy Excel Columns A-F,J-L, Q, S; create new dataframe
## Variables in this set are weird. 
## 1. You need to select columns you want to keep/drop from `Shaw Full Profiles`
##        You can drop the calculated columns that you will calculate yourself (ccontent, etc.)
## 2. Select columns you want from `Shaw (only sites)`
## 3. Join the two dataframes
## 4. Calculate carbon stock columns and sum in summary tables

```{r new-dataframe-with-select-variables, eval=T, echo=FALSE, message=FALSE, warning=FALSE}

``` 

## Create an index column specifiying if a bulk density value is gap-filled or not

Create a vector of gap-filled values

```{r bulk-density-gap-filled-values, echo=FALSE}

```

Create the `bd_fill` index column

```{r create-gap-filled-index, echo=FALSE}

```

## Extract soil horizon from `horizon` and call it `horizon_simple`
```{r create-horizon-simple, echo=FALSE}

```
## Calculate carbon content for each horizon

```{r mutate-cf-na-to-0, echo=FALSE, eval=FALSE}

```


## data_subset <- data_subset %>%
## mutate(ccontent = (cconc/100*bulk_density)*
##                  (100*100*total_depth)*
##                ((100 - cf)/100))
##data_subset


## Create carbon_stock and carbon_stock_1M variables
```{r create-shaw-full-profiles-carbon_stock-variable}

```
```{r create-shaw-full-profiles-carbon_stock_1M-variable}

```


## Save the tidy dataset as a .csv

```{r, write-csv, echo=FALSE}

```

## Create summary tables for `carbon_stock`, `carbon_stock_1m`

```{r carbon_stock-summary-table, echo=FALSE}

```

```{r carbon_stock_1m-summary-table, echo=FALSE}

```