---
title: "Calvert_final"
output: html_document
date: '2022-04-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
# install.packages("tindyverse")
# install.packages("readxl")
library("readxl")
library("tidyverse")
library("dplyr")
library("openintro")
library("skimr")
library("janitor")
library("psych")
```


## R Markdown
## Explore Calvert_Final Batter data, clean col names, convert to lowercase, tidy data classes
```{r}
library(readxl)
X3_Calvert_FINAL <- read_excel("3 Calvert FINAL.xlsx")
View(X3_Calvert_FINAL)
```


### Add another line in your mutate function to calculate horizon depth >- 100 cm

```{r read-in-krana-better-banner-data, warning = F}
calvert_dataset <- X3_Calvert_FINAL %>%  
  janitor::clean_names() %>% 
  rename_with(tolower) %>%
  mutate(bulk_density = as.numeric(bulk_density),
         depth1 = as.numeric(depth1),
         depth2 = as.numeric(depth2),
         cconc = as.numeric(cconc),
         total_depth=depth2-depth1,
         depth_100 = ifelse(depth2<=-100,
                            abs(depth1-(-100)),
                            abs(depth2-depth1)))
glimpse(calvert_dataset)
```
## Retain Tidy Excel Columns A-F,J-L, Q, S

```{r new-dataframe-with-select-variables}
calvert_dataset <- calvert_dataset %>% 
  select(source, 
         id, 
         order, 
         lat, 
         lon, 
         latlon_q, 
         horizon, 
         depth1, 
         depth2, 
         total_depth, 
         bulk_density, 
         cf, 
         cconc)
glimpse(calvert_dataset)
``` 
## Create an index column specifiying if a value is gap-filled or not
## All bulk_density is gapfilled in kranabetter-batter dataset

Create a vector of gap-filled values

```{r gap-filled-values}
bd_fill_values <- calvert_dataset %>%
  select(bulk_density) %>%
  filter(!is.na(bulk_density)) %>% 
  distinct() %>% 
  pull() 
bd_fill_values
```

## mutate cconc gapfilled columns by making function (see note card)

```{r}
calvert_dataset <- calvert_dataset %>%
mutate(slope = (cconc/total_depth),
       cconc_y = (cconc-(slope*.5*total_depth))
       )
calvert_dataset
```


## Extract first soil horizon from 'horizon', name column 'horizon simple'
```{r create-horizon-simple}
calvert_dataset$horizon_simple = substr(calvert_dataset$horizon,1,1)
calvert_dataset
```
## Calculate carbon content for each horizon in kranabetter dataset

```{r mutate-na-in-cf-to-0}
calvert_dataset <- calvert_dataset %>%
 mutate(cf = ifelse(is.na(cf), 0, cf),
        cf_1_values = as.numeric(cf))
calvert_dataset
```

```{r calculate-ccontent-from-function-for-kranabetter-data}
calvert_dataset <- calvert_dataset %>%
mutate(ccontent = (cconc/100*bulk_density)*(100*100*total_depth)*((100 - cf)/100))
calvert_dataset
```

## Save the tidy dataset as a .csv

```{r}
write.csv(kranabetter_data_subset, 'Kranabetter_Banner_Tidy.csv')
```

## Create summary tables for `carbon_stock`, `carbon_stock_1m`

```{r carbon_stock-summary-table}
describeBy(calvert_dataset, 
           group=calvert_dataset$carbon_stock
           )
```

```{r carbon_stock_1m-summary-table}
describeBy(calvert_dataset, 
           group=calvert_dataset$carbon_stock_1m
           )
```

