---
title: "DAmore_2002_Final"
author: "Desira Miller"
date: '2022-06-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r load-packages, include = FALSE}
# install.packages("tindyverse")
# install.packages("readxl")
library("readxl")
library("tidyverse")
library("dplyr")
library("openintro")
library("skimr")
library("janitor")
library("psych")
library("stringr")
```

## Explore data, clean col names, convert to lowercase
### Add another line in your mutate function to calculate horizon depth >- 100 cm

```{r read-in-darmore-2002-final-data, warning=F}
data_darmore <- read_excel("7 D'Armore 2002 FINAL.xlsx") %>%  
  janitor::clean_names() %>% 
  rename_with(tolower) %>%
  mutate(bulk_density = as.numeric(bulk_density),
         bulk_density = ifelse(is.na(bulk_density),
                              0, bulk_density),
         depth1 = as.numeric(depth1),
         depth2 = as.numeric(depth2),
         cconc = as.numeric(cconc),
         cconc = ifelse(is.na(cconc),
                             0, cconc),
         cf = as.numeric(mincont),
         cf = ifelse(is.na(cf),
                              0, cf),
         ccontent_1m = as.numeric(ccontent_1m),
         ccontent_1m = ifelse(is.na(ccontent_1m),
                             0, ccontent_1m),
         total_depth=depth2-depth1,
         depth_100 = ifelse(depth2<=-100,
                            abs(depth1-(-100)),
                            abs(depth2-depth1))
  )
glimpse(data_darmore)
```

## Coerce NAs in bulk_denisty and cconc to 0
```{r mutate-NAs-to-0}
data_darmore <- data_darmore %>%
 mutate(bulk_density = ifelse(is.na(bulk_density),
                              0, bulk_density),
                cconc = ifelse(is.na(cconc),
                             0, cconc),
                  cf = ifelse(is.na(cf),
                              0, cf),
                ccontent_1m = ifelse(is.na(ccontent_1m),
                              0, ccontent_1m))
glimpse(data_darmore)
```

## Retain Relevant Tidy Excel Columns; create new dataframe

```{r new-dataframe-with-select-variables}
subset_darmore <- data_darmore %>% 
  select(source, 
         id, 
         order, 
         lat, 
         lon, 
         latlon_q, 
         horizon, 
         depth1, 
         depth2, 
         total_depth, 
         depth_100,
         bulk_density, 
         cconc,
         cf,
         ccontent_1m)
glimpse(subset_darmore)
``` 

## Create an index column specifiying if a bulk density value is gap-filled or not
## Create an index column specifiying if a value is gap-filled or not
## All bulk_density is gapfilled in D'armore dataset

```{r gap-filled-values}
bd_fill_values_darmore <- subset_darmore %>%
  select(bulk_density) %>%
  filter(!is.na(bulk_density)) %>% 
  distinct() %>% 
  pull() 
bd_fill_values_darmore
```

Create the `bd_fill` index column

```{r create-gap-filled-index}
subset_darmore <- subset_darmore %>% 
  mutate(bd_filled = ifelse(bulk_density %in% bd_fill_values_darmore, "yes", "no")) 
subset_darmore
```

## Extract soil horizon from 'horizon' and call it 'horizon_simple'
```{r create-horizon-simple}
subset_darmore$horizon_simple = substr(subset_darmore$horizon,1,1)
subset_darmore
```
## Calculate carbon content for each horizon
```{r mutate-cf-na-to-0}
subset_darmore <- subset_darmore %>%
 mutate(cf = ifelse(is.na(cf), 0, cf),
        cf_1_values = as.numeric(cf))
subset_darmore
```

```{r calculate-ccontent-from-function}
subset_darmore <- subset_darmore %>%
mutate(ccontent = (cconc/100*bulk_density)*
                  (100*100*total_depth)*
                  ((100 - cf)/100))
subset_darmore
```
## Create carbon_stock and carbon_stock_1M variables
```{r create-carbon_stock-variable}
subset_darmore <- subset_darmore %>%
mutate(carbon_stock = (cconc*total_depth*bulk_density))
subset_darmore
```
```{r create-carbon_stock_1M-variable}
subset_darmore <- subset_darmore %>%
mutate(carbon_stock_1m = (ccontent_1m * depth_100 * bulk_density))
subset_darmore
```


## Save the tidy dataset as a .csv

```{r}
write.csv(subset_darmore, 'DArmore_2002_Final_Tidy.csv')
```

## Create summary tables for `carbon_stock`, `carbon_stock_1m`

```{r carbon_stock-summary-table}
describeBy(subset_darmore, group=subset_darmore$carbon_stock)
```

```{r carbon_stock_1m-summary-table}
describeBy(subset_darmore, group=subset_darmore$carbon_stock_1m)
```