---
title: "Dryad Soil Data Wrangling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
# install.packages("tindyverse")
# install.packages("readxl")
library("readxl")
library("tidyverse")
library("dplyr")
library("openintro")
library("skimr")
library("janitor")
library("psych")
library("stringr")
```


## R Markdown

## Please email Desira Miller at desiraaa@gmail.com if you have any questions pertaining to the code

## Explore data, clean col names, convert to lowercase, 
### Add another line in your mutate function to calculate horizon depth >- 100 cm

```{r read-in-sandborn_lewis_data, warning = F}
sandborn_lewis_data <- read_excel("1 Sandborn Lewis data.xlsx") %>%  
  janitor::clean_names() %>%  ##replaces spaces in variable names with '_'
  rename_with(tolower) %>% ##lowers all uppercase letters to lowercase
  mutate(bulk_density = as.numeric(bulk_density), ## wrangle bulk_density, depth1, depth2, cconc, ccontent_1m to numeric variables
         depth1 = as.numeric(depth1),
         depth2 = as.numeric(depth2),
         cconc = as.numeric(cconc),
         ccontent_1m = as.numeric(ccontent_1m),
         total_depth=depth2-depth1, ## create total_depth variable by subtracting depth2 from depth1 
         depth_100 = ifelse(depth2<=-100, ## create depth_100 variable 
                            abs(depth1-(-100)),
                            abs(depth2-depth1))
  )
glimpse(sandborn_lewis_data)
```

## Retain Tidy Excel Columns A-F,J-L, Q, S; create new dataframe
## These are generally the main variables retained from each dataset to create a new subset

```{r new-dataframe-with-select-variables}
sandborn_lewis_data_subset <- sandborn_lewis_data %>% 
  select(source, 
         id, 
         order, 
         lat, 
         lon, 
         latlon_q, 
         horizon, 
         depth1, 
         depth2, 
         total_depth,
         bulk_density, 
         cf, 
         cconc,
         ccontent_1m,
         depth_100)
glimpse(sandborn_lewis_data_subset)
``` 

## Create an index column specifiying if a bulk density value is gap-filled or not

Create a vector of gap-filled values

```{r bulk-density-gap-filled-values}
bd_fill_values <- sandborn_lewis_data_subset %>%
  select(bulk_density) %>%
  slice(1:3, 8:10, 16:28, 34:39, 46:146) %>%
  filter(!is.na(bulk_density)) %>% 
  distinct() %>% 
  pull() 
bd_fill_values ## bd_fill_values is a new column that contains gap filled values from bulk density variable. if column is not gap filled then the value equals 0
```

Create the `bd_fill` index column

```{r create-gap-filled-index}
sandborn_lewis_data_subset <- sandborn_lewis_data_subset %>% 
  mutate(bd_filled = ifelse(bulk_density %in% bd_fill_values, "yes", "no")) ## bd_filled is a new variable indicating whether or not bulk_density variable is gap filled 
sandborn_lewis_data_subset
```

Identified a weird row for a single profile `id` missing all important variables. 

Drop the weird row `id` = `BC09-07` where `horizon` = `Cg`

```{r remove-weird-row}
sandborn_lewis_data_subset <- sandborn_lewis_data_subset %>% 
  filter(id != "BC09-06" | horizon != "Cg")
```

## Extract first letter from soil horizon from `horizon` and call it `horizon_simple`
```{r create-horizon-simple}
sandborn_lewis_data_subset$horizon_simple = substr(sandborn_lewis_data_subset$horizon,1,1)
sandborn_lewis_data_subset
```
## Calculate carbon content for each horizon
## Wrangle NA values to 0

```{r mutate-cf-na-to-0}
sandborn_lewis_data_subset <- sandborn_lewis_data_subset %>%
 mutate(cf = ifelse(is.na(cf), 0, cf),
        cf_1_values = as.numeric(cf))
sandborn_lewis_data_subset
```

```{r calculate-ccontent-from-function}
sandborn_lewis_data_subset <- sandborn_lewis_data_subset %>%
mutate(ccontent = (cconc/100*bulk_density)*
                  (100*100*total_depth)*
                  ((100 - cf)/100)) ## ccontent = carbon content created from following function
sandborn_lewis_data_subset
```
## Create carbon_stock and carbon_stock_1M variables
```{r create-carbon_stock-variable}
sandborn_lewis_data_subset <- sandborn_lewis_data_subset %>%
mutate(carbon_stock = (cconc*total_depth*bulk_density))
sandborn_lewis_data_subset
```
```{r create-carbon_stock_1M-variable}
sandborn_lewis_data_subset <- sandborn_lewis_data_subset %>%
mutate(carbon_stock_1m = (ccontent_1m * depth_100 * bulk_density))
sandborn_lewis_data_subset
```

## Save the tidy dataset as a .csv

```{r}
write.csv(sandborn_lewis_data_subset, 'Sandborn_Lewis_Tidy.csv')
```

## Create summary tables for `carbon_stock`, `carbon_stock_1m`

```{r carbon_stock-summary-table}
describeBy(sandborn_lewis_data_subset, groupby=sandborn_lewis_data_subset$carbon_stock)
```

```{r carbon_stock_1m-summary-table}
describeBy(sandborn_lewis_data_subset, group=sandborn_lewis_data_subset$carbon_stock_1m)
```
