---
title: "Shaw_Only_Sites"
output: html_document
date: "`2022-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
# install.packages("tindyverse")
# install.packages("readxl")
library("readxl")
library("tidyverse")
library("dplyr")
library("openintro")
library("skimr")
library("janitor")
library("psych")
```


## R Markdown
## Explore data, clean col names, convert to lowercase, tidy data classes

```{r read-in-shaw-only-sites, eval=T, echo=FALSE, message=FALSE, warning=FALSE}
shaw_only_sites_data <- read_excel("5 Shaw (only sites).xlsx") %>%
  janitor::clean_names() %>% 
  rename_with(tolower)
shaw_only_sites_data
```


### Add another line in your mutate function to calculate horizon depth >- 100 cm


## Retain Relevant Tidy Excel Columns 
## Retaining these idenitites are not needed


## Rows 22-40 and 46-76 are not used due to low lat/lon quality in original dataset


## Soil horizons are written as their full name in original dataset. Extract simple horizon names somehow?

## Calculate carbon content for each horizon in shaw_only_sites dataset

## calculate carbon content


## Save the tidy dataset as a .csv


## Create summary tables for `carbon_stock`, `carbon_stock_1m`


